---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE} 
library(dplyr)
library(tidyr)
library(readr)
```

message = FALSE, warning = FALSE. will not list these errors in knitted document

using just the URL in read.csv can read directly from the download link instead of downloading. 
```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

readr package is read_csv is a bit smarter for guessing column types and a bit more efficient when reading in larger files

##Split-Apply-Combine
```{r}
head(catch_original)
```


```{r}
catch_data <- catch_original %>% 
  #select(Region, Year, Chinook, Coho, Pink, Chum, Sockeye)
#head(catch_data)

select(-All, -notesRegCode)

glimpse(catch_data)
summary(catch_data)
```

- change Chinook from character to numeric type.
mutate can do this
NAs introduced by coercion: didn't know what to do with the value
```{r}
catch_clean <- catch_data %>% mutate(Chinook = as.numeric(Chinook) )
```
###finding spurious characters in Chinook column
which returns numbers for which a statement is true
and then go find the row
go check the original data (perhaps an I instead of a 1). 
```{r}
i <- which (is.na(catch_clean$Chinook))
i
catch_original[i,]
```
use mutate
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I","1", Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))
```
##reshaping data
c function specifies a vector

rename() can rename columns
```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region,Year), names_to = "Species",
               values_to = "Catch") %>% 
  rename(catch_thousands = Catch) %>%
  mutate(catch = catch_thousands*1000) %>%
  select(-catch_thousands)

head(catch_long)
```

```{r}
catch_wide <- catch_long %>%
  pivot_wider(names_from = Region, values_from = catch)
```

if the column name already exists, it can be without "
if you are making a new column add "

##split apply combine

-calculate mean salmon catch by region
(split along region varaible and apply mean function by catch)

filter- grabs rows
group by - graps columns

```{r}
mean_region <- catch_long %>% 
  filter(Species == "Sockeye") %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(catch),
            n = n(),
            .groups = "drop")

head(mean_region)
```

 %>% rename(Region, code)
 can use this to rename the code

```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

use c for combine column
```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)

nrow(catch_joined) == nrow(catch_long)
```

separate and unite are also helpful. 

##Joins

%>% 