---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

### Smaller Header

Write some text here. **Using the shift-8 star keys will make text bold**.

create a list with asterick

-   apple
-   banana
-   kiwi
-   need a blank line to start a list

To complete the task, I am writing in *italic*.

A numbered list below:

1.  list
2.  list
3.  list

## I think sub header is Header 2

#### subheader is the 4th level

insert citation
using the far right diagram to enter a citation, visual markdown

add a biobilography section, references.bib is in the directory. bibtech document
the reference to the citation will be included in the line. 

markdown table synteax is annoying, so the insert selection is helpful. 


how to add Rcode into the document. 

```{r}
x <- 3+2

y <- 2+7
```
```{r}

```


##Read in data

```{r}
bg_chem <- read.csv("../Data/BGchem2008data.csv")
```

```{r}
read.csv(" ../Data/")

read.csv("Data/BGchem2008data.csv")

```


## Analysis

## Plots


**Practicing uploading RMarkdown to Git**

##This is a header

It doesn't seem to be working. 

Is this working?

